---
title: "Impact of Tech Use on Mental Health"
author: "Anurag Mallick"
date: "2025-05-03"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
fontsize: 12pt
classoption: twoside
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(dplyr)
```

## Effect on Mental Health

We study mental health through the NSDUH survey which has certain indicator variables to help us gauge the effect on mental health.

-   Feeling of Nervousness

-   Feeling of Hopelessness

-   Feeling of Restlessness


## Psychological Distress

We start with exploring the nervousness, hopelessness and restlessness for a period of 2002-2008.
```{r, warning=FALSE, message=FALSE, echo=FALSE}
psych_summ_ovrl <- read_csv("health_eda/variable_eda/psych_summ_ovrl.csv", show_col_types = FALSE)
psych_summ_age_ovrl <- read_csv("health_eda/variable_eda/psych_summ_age_ovrl.csv", show_col_types = FALSE)
psych_summ_inc_ovrl <- read_csv("health_eda/variable_eda/psych_summ_income_ovrl.csv", show_col_types = FALSE)
psych_summ_edu_ovrl <- read_csv("health_eda/variable_eda/psych_summ_edu_ovrl.csv", show_col_types = FALSE)
```

## Nervousness
```{r}
psych_summ_ovrl |> 
  mutate(Nervous_123_percent=(Nervous_1+Nervous_2+Nervous_3)/num_people,
         Nervous_1_percent=Nervous_1/num_people,
         Nervous_any_percent=Nervous_any/num_people) |> 
  pivot_longer(cols=c(Nervous_123_percent,Nervous_1_percent, Nervous_any_percent), values_to = "Nervous_Values", names_to = "Nervous_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Nervous_Values, color = Nervous_type)) +
  geom_line(size=1.5) +
  
  
```

### Different Age Groups
```{r}
psych_summ_age_ovrl |> 
  mutate(Nervous_123_percent=(Nervous_1+Nervous_2+Nervous_3)/num_people,
         Nervous_1_percent=Nervous_1/num_people,
         Nervous_any_percent=Nervous_any/num_people) |> 
  pivot_longer(cols=c(Nervous_123_percent,Nervous_1_percent, Nervous_any_percent), values_to = "Nervous_Values", names_to = "Nervous_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Nervous_Values, group = CATAG7, color=CATAG7)) +
  geom_line(size=1.5)+
  facet_wrap(~ Nervous_type)
```


## Hopelessness
```{r}
psych_summ_ovrl |> 
  mutate(Hopeless_123_percent=(Hopeless_1+Hopeless_2+Hopeless_3)/num_people,
         Hopeless_1_percent=Hopeless_1/num_people,
         Hopeless_any_percent=Hopeless_any/num_people) |> 
  pivot_longer(cols=c(Hopeless_123_percent,Hopeless_1_percent, Hopeless_any_percent), values_to = "Hopeless_Values", names_to = "Hopeless_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Hopeless_Values, color = Hopeless_type)) +
  geom_line(size=1.5)
```


### Different Age Groups
```{r}
psych_summ_age_ovrl |> 
  mutate(Hopeless_123_percent=(Hopeless_1+Hopeless_2+Hopeless_3)/num_people,
         Hopeless_1_percent=Hopeless_1/num_people,
         Hopeless_any_percent=Hopeless_any/num_people) |> 
  pivot_longer(cols=c(Hopeless_123_percent,Hopeless_1_percent, Hopeless_any_percent), values_to = "Hopeless_Values", names_to = "Hopeless_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Hopeless_Values,  group = CATAG7, color=CATAG7)) +
  geom_line(size=1.5)+
  facet_wrap(~ Hopeless_type)
```
