---
title: "Visualizing Mental Health"
author: "Anurag Mallick"
date: "2025-05-01"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
fontsize: 12pt
classoption: twoside
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(dplyr)
```

## Effect on Mental Health

We study mental health through the NSDUH survey which has certain indicator variables to help us gauge the effect on mental health.

-   Feeling of Nervousness

-   Feeling of Hopelessness

-   Feeling of Restlessness

## Psychological Distress

We start with exploring the nervousness, hopelessness and restlessness for a period of 2002-2008.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
psych_summ_ovrl <- read_csv("health_eda/variable_eda/psych_summ_ovrl.csv", show_col_types = FALSE)
psych_summ_age_ovrl <- read_csv("health_eda/variable_eda/psych_summ_age_ovrl.csv", show_col_types = FALSE)
psych_summ_inc_ovrl <- read_csv("health_eda/variable_eda/psych_summ_income_ovrl.csv", show_col_types = FALSE)
psych_summ_edu_ovrl <- read_csv("health_eda/variable_eda/psych_summ_edu_ovrl.csv", show_col_types = FALSE)
```

```{r}
# This has each strata grouped by Gender, Age, Income and Education Level
psych_summ_all_grp_ovrl <- read_csv("health_eda/variable_eda/psych_summ_all_grp_ovrl.csv", show_col_types = FALSE)
```

## Nervousness

Nervousness number 1,2,3,4 shows different level of nervousness for the past 12 months.

1 = All of the time 
2 = Most of the time 
3 = Some of the time 
4 = A little of the time 

### Overall for different years
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |> 
  select(-c(IRSEX, CATAG7,INCOME, EDUCCAT2)) |> 
  group_by(YEAR) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Nervous_123_percent=(Nervous_1+Nervous_2+Nervous_3)/num_people,
         Nervous_1_percent=Nervous_1/num_people,
         Nervous_12_percent=(Nervous_1+Nervous_2)/num_people,
         Nervous_any_percent=Nervous_any/num_people) |> 
  pivot_longer(cols=c(Nervous_1_percent, Nervous_12_percent, Nervous_123_percent,Nervous_any_percent), 
               values_to = "Nervous_Values", names_to = "Nervous_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Nervous_Values, color = Nervous_type)) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Nervous") +
  ggtitle("Percentage Nervous for different years") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(
      values = c("Nervous_1_percent" = "brown","Nervous_12_percent" = "maroon","Nervous_123_percent" = "orange", 
                 "Nervous_any_percent" = "red"),
      labels = c("All Times", "All or Most Times", "Little Time", "Atleast one time"),name = "Nervous Category")
```

### Different Genders
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |> 
  select(-c(CATAG7,INCOME, EDUCCAT2)) |> 
  group_by(YEAR, IRSEX) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Nervous_1_percent=Nervous_1/num_people,
         Nervous_any_percent=Nervous_any/num_people) |> 
  pivot_longer(cols=c(Nervous_1_percent, Nervous_any_percent),
               values_to = "Nervous_Values", names_to = "Nervous_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Nervous_Values, color = as.factor(IRSEX), group = as.factor(IRSEX))) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Nervous") +
  ggtitle("Percentage Nervous by gender") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(
      values = c("1" = "orange", "2" = "cyan"),
      labels = c("1" = "Male", "2" = "Female"),name = "Gender Category")  +
  facet_wrap(~Nervous_type)
```


### Different Age Groups
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |>
  select(-c(IRSEX,INCOME, EDUCCAT2)) |> 
  group_by(YEAR, CATAG7) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Nervous_1_percent=Nervous_1/num_people,
         Nervous_any_percent=Nervous_any/num_people) |> 
  pivot_longer(cols=c(Nervous_1_percent, Nervous_any_percent),
               values_to = "Nervous_Values", names_to = "Nervous_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Nervous_Values, color = as.factor(CATAG7), group = as.factor(CATAG7))) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Nervous") +
  ggtitle("Percentage Nervous by age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Nervous_type)
```


### Different Income Groups
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |>
  select(-c(IRSEX, CATAG7, EDUCCAT2)) |> 
  group_by(YEAR, INCOME) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Nervous_1_percent=Nervous_1/num_people,
         Nervous_any_percent=Nervous_any/num_people) |> 
  pivot_longer(cols=c(Nervous_1_percent, Nervous_any_percent),
               values_to = "Nervous_Values", names_to = "Nervous_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Nervous_Values, color = as.factor(INCOME), group = as.factor(INCOME))) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Nervous") +
  ggtitle("Percentage Nervous by income") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Nervous_type) + 
  scale_colour_manual(name = "Income Group",
    values = c("1" = "lightblue","2" = "orange","3" = "darkgrey","4" = "maroon"),
    labels = c("1" = "Less than $20,000","2" = "$20,000 - $49,999","3" = "$50,000 - $74,999","4" = "$75,000 or More"))
```




## Hopelessness

### Overall for different years
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |> 
  select(-c(IRSEX, CATAG7,INCOME, EDUCCAT2)) |> 
  group_by(YEAR) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Hopeless_123_percent=(Hopeless_1+Hopeless_2+Hopeless_3)/num_people,
         Hopeless_1_percent=Hopeless_1/num_people,
         Hopeless_12_percent=(Hopeless_1+Hopeless_2)/num_people,
         Hopeless_any_percent=Hopeless_any/num_people) |> 
  pivot_longer(cols=c(Hopeless_1_percent, Hopeless_12_percent, Hopeless_123_percent,Hopeless_any_percent), 
               values_to = "Hopeless_Values", names_to = "Hopeless_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Hopeless_Values, color = Hopeless_type)) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Hopeless") +
  ggtitle("Percentage Hopeless for different years") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(
      values = c("Hopeless_1_percent" = "brown","Hopeless_12_percent" = "maroon","Hopeless_123_percent" = "orange", 
                 "Hopeless_any_percent" = "red"),
      labels = c("All Times", "All or Most Times", "Little Time", "Atleast one time"),name = "Hopeless Category")
```



### Different Age Groups
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |>
  select(-c(IRSEX,INCOME, EDUCCAT2)) |> 
  group_by(YEAR, CATAG7) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Hopeless_1_percent=Hopeless_1/num_people,
         Hopeless_any_percent=Hopeless_any/num_people) |> 
  pivot_longer(cols=c(Hopeless_1_percent, Hopeless_any_percent),
               values_to = "Hopeless_Values", names_to = "Hopeless_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Hopeless_Values, color = as.factor(CATAG7), group = as.factor(CATAG7))) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Hopeless") +
  ggtitle("Percentage Hopeless by age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Hopeless_type)
```





## Restlessness

### Overall for different years
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |> 
  select(-c(IRSEX, CATAG7,INCOME, EDUCCAT2)) |> 
  group_by(YEAR) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Restless_123_percent=(Restless_1+Restless_2+Restless_3)/num_people,
         Restless_1_percent=Restless_1/num_people,
         Restless_12_percent=(Restless_1+Restless_2)/num_people,
         Restless_any_percent=Restless_any/num_people) |> 
  pivot_longer(cols=c(Restless_1_percent, Restless_12_percent, Restless_123_percent,Restless_any_percent), 
               values_to = "Restless_Values", names_to = "Restless_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Restless_Values, color = Restless_type)) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Restless") +
  ggtitle("Percentage Restless for different years") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(
      values = c("Restless_1_percent" = "brown","Restless_12_percent" = "maroon","Restless_123_percent" = "orange", 
                 "Restless_any_percent" = "red"),
      labels = c("All Times", "All or Most Times", "Little Time", "Atleast one time"),name = "Restless Category")
```



### Different Age Groups
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7>=4) |>
  select(-c(IRSEX,INCOME, EDUCCAT2)) |> 
  group_by(YEAR, CATAG7) |> 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(Restless_1_percent=Restless_1/num_people,
         Restless_any_percent=Restless_any/num_people) |> 
  pivot_longer(cols=c(Restless_1_percent, Restless_any_percent),
               values_to = "Restless_Values", names_to = "Restless_type") |> 
  ggplot(mapping=aes(x=YEAR, y=Restless_Values, color = as.factor(CATAG7), group = as.factor(CATAG7))) +
  geom_line(size=1.5)+
  xlab("Year") +
  ylab("% Restless") +
  ggtitle("Percentage Restless by age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~Restless_type)
```


## For all emotions

### By Income Group
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7 >= 4) |>
  select(-c(IRSEX, CATAG7, EDUCCAT2)) |>
  group_by(YEAR, INCOME) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    Nervous_1_percent = Nervous_1 / num_people,
    Nervous_any_percent = Nervous_any / num_people,
    Hopeless_1_percent = Hopeless_1 / num_people,
    Hopeless_any_percent = Hopeless_any / num_people,
    Restless_1_percent = Restless_1 / num_people,
    Restless_any_percent = Restless_any / num_people
  ) |>
  pivot_longer(
    cols = ends_with("_percent"),
    names_to = c("Distress", "Period"),
    names_pattern = "(.*)_(1|any)_percent",
    values_to = "Percent"
  ) |>
  mutate(
    Distress = case_when(
      grepl("Nervous", Distress) ~ "Nervous",
      grepl("Hopeless", Distress) ~ "Hopeless",
      grepl("Restless", Distress) ~ "Restless"
    ),
    Period = recode(Period, `1` = "All the time", `any` = "Any time")
  ) |>
  ggplot(aes(x = YEAR, y = Percent, color = as.factor(INCOME), group = as.factor(INCOME))) +
  geom_line(size = 1.2) +
  facet_grid(Period ~ Distress) +
  xlab("Year") +
  ylab("Percentage of People in Distress") +
  ggtitle("Psychological Distress by Income Group") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_manual(name = "Income Group",
    values = c("1" = "lightblue","2" = "orange","3" = "darkgrey","4" = "maroon"),
    labels = c("1" = "Less than $20,000","2" = "$20,000 - $49,999","3" = "$50,000 - $74,999","4" = "$75,000 or More"))
  
```


### By Age Group
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7 >= 4) |>
  select(-c(IRSEX, INCOME, EDUCCAT2)) |>
  group_by(YEAR, CATAG7) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    Nervous_1_percent = Nervous_1 / num_people,
    Nervous_any_percent = Nervous_any / num_people,
    Hopeless_1_percent = Hopeless_1 / num_people,
    Hopeless_any_percent = Hopeless_any / num_people,
    Restless_1_percent = Restless_1 / num_people,
    Restless_any_percent = Restless_any / num_people
  ) |>
  pivot_longer(
    cols = ends_with("_percent"),
    names_to = c("Distress", "Period"),
    names_pattern = "(.*)_(1|any)_percent",
    values_to = "Percent"
  ) |>
  mutate(
    Distress = case_when(
      grepl("Nervous", Distress) ~ "Nervous",
      grepl("Hopeless", Distress) ~ "Hopeless",
      grepl("Restless", Distress) ~ "Restless"
    ),
    Period = recode(Period, `1` = "All the time", `any` = "Any time")
  ) |>
  ggplot(aes(x = YEAR, y = Percent, color = as.factor(CATAG7), group = as.factor(CATAG7))) +
  geom_line(size = 1.2) +
  facet_grid(Period ~ Distress) +
  xlab("Year") +
  ylab("Percentage of People in Distress") +
  ggtitle("Psychological Distress by Age Group") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_manual(name = "Age Group",
    values = c("4" = "lightblue","5" = "orange","6" = "darkgrey","7" = "maroon"),
    labels = c("4" = "18-20 Years Old","5" = "21-25 Years Old","6" = "26-34 Years Old","7" = ">35 Years Old"))
  
```


### By Education Group
```{r fig.width=10}
psych_summ_all_grp_ovrl |>
  filter(CATAG7 >= 4) |>
  select(-c(IRSEX, INCOME, CATAG7)) |>
  group_by(YEAR, EDUCCAT2) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    Nervous_1_percent = Nervous_1 / num_people,
    Nervous_any_percent = Nervous_any / num_people,
    Hopeless_1_percent = Hopeless_1 / num_people,
    Hopeless_any_percent = Hopeless_any / num_people,
    Restless_1_percent = Restless_1 / num_people,
    Restless_any_percent = Restless_any / num_people
  ) |>
  pivot_longer(
    cols = ends_with("_percent"),
    names_to = c("Distress", "Period"),
    names_pattern = "(.*)_(1|any)_percent",
    values_to = "Percent"
  ) |>
  mutate(
    Distress = case_when(
      grepl("Nervous", Distress) ~ "Nervous",
      grepl("Hopeless", Distress) ~ "Hopeless",
      grepl("Restless", Distress) ~ "Restless"
    ),
    Period = recode(Period, `1` = "All the time", `any` = "Any time")
  ) |>
  ggplot(aes(x = YEAR, y = Percent, color = as.factor(EDUCCAT2), group = as.factor(EDUCCAT2))) +
  geom_line(size = 1.2) +
  facet_grid(Period ~ Distress) +
  xlab("Year") +
  ylab("Percentage of People in Distress") +
  ggtitle("Psychological Distress by Education Group") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_manual(name = "Education Group",
    values = c("1" = "lightblue","2" = "orange","3" = "darkgrey","4" = "maroon"),
    labels = c("1" = "Less than high school","2" = "High School","3" = "Some college","4" = "College Graduate"))
  
```
