---
title: "Integrated Visualization"
author: "Anurag Mallick, Sanjana Umesh Sawant"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(ggplot2)
library(scales)
```



## Tech Data

```{r}
tech_use_df <- read_csv("tech_eda/results/tech_use_summary.csv")
net_use_df <- read_csv("tech_eda/results/net_use_summary.csv")
```

```{r}
tech_net_use_df <- tech_use_df |> inner_join(net_use_df)
```

```{r}
# get the tech use and net use
tech_net_use_df <- tech_net_use_df |> 
  mutate(tech_use = as.integer(avg_tech_use*n/100), 
         net_use = as.integer(avg_net_use*n/100))


tech_net_use_df |> head(5)
```


## Mental Health Data
```{r}
mental_health_df <- read_csv("health_eda/variable_eda/psych_summ_all_grp_ovrl.csv", show_col_types = FALSE)
```

```{r}
mental_health_df |> head(5)
```



## Integrating on attributes

In tech data, we have the attributes of gender, age and income while in health data we have education also. To merge we consider the combination of the three common demographics - gender, age and income. 


```{r}
mental_health_df_grp <- mental_health_df |>
  select(-EDUCCAT2) |> 
  group_by(YEAR, IRSEX, CATAG7, INCOME) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |> 
  ungroup() |> 
  rename(c(Gender=IRSEX, Age=CATAG7, Income=INCOME))

tech_net_use_df <- tech_net_use_df |> 
  rename(c(YEAR=hryear4, Gender=pesex, Age=prtage, Income=hufaminc))
```

```{r}
mental_health_df_grp |> head(2)
```

```{r}
tech_net_use_df |> head(2)
```

```{r}
tech_mental_df <- tech_net_use_df |> 
  full_join(mental_health_df_grp, by=c("YEAR", "Age", "Income", "Gender")) 
```

```{r}
tech_mental_df |> head(5)
```


```{r}
tech_mental_df |> filter(YEAR==2009)
```


### Storing tech_mental_data
```{r}
#write_csv(tech_mental_df,"final_data/final_tech_health.csv")
```



## Visualization

### Getting Data
```{r}
tech_mental_df <- read_csv("final_data/final_tech_health.csv")
```

### Tech Users and Psych Distress in different years
```{r fig.width=10}
p_1 <- tech_mental_df |>
  group_by(YEAR) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2027) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Nervous_1_percent = (Nervous_1 / num_people)
  ) |>
  select(YEAR, Nervous_1_percent, net_use_percent) |> 
  pivot_longer(
    cols = c("Nervous_1_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  na.omit() |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("Percentage of People Nervous  All Time") +
  ggtitle("Psychological Nervousness") +
  theme(plot.title = element_text(hjust = 0.5))

ggsave("plots/p_tech_psych.png", plot = p_1, width = 10, height = 10)

p_1
```


### Tech Users and Psych - Nervous in diff Genders
```{r fig.width=10}
p_2 <- tech_mental_df |>
  group_by(YEAR, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Nervous_1_percent = (Nervous_1 / num_people),
    Nervous_any_percent = (Nervous_any / num_people)
  ) |>
  select(YEAR, Gender, Nervous_1_percent,Nervous_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Nervous_1_percent", "Nervous_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_wrap(~Gender_Str) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Nervous ") +
  ggtitle("Internet Use and Psychological Nervousness") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Nervous_1_percent" = "lightblue","Nervous_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Nervous_1_percent" = "Nervous All Time","Nervous_any_percent" = "Nervous Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_nervous_gender.png", plot = p_2, width = 10, height = 10)

p_2
```


### Tech Users and Psych - Hopeless in diff Genders
```{r fig.width=10}
p_3 <- tech_mental_df |>
  group_by(YEAR, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Hopeless_1_percent = (Hopeless_1 / num_people),
    Hopeless_any_percent = (Hopeless_any / num_people)
  ) |>
  select(YEAR, Gender, Hopeless_1_percent,Hopeless_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Hopeless_1_percent", "Hopeless_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_wrap(~Gender_Str) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Hopeless") +
  ggtitle("Internet Use and Psychological Hopelessness") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Hopeless_1_percent" = "lightblue","Hopeless_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Hopeless_1_percent" = "Hopeless All Time","Hopeless_any_percent" = "Hopeless Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_hopeless_gender.png", plot = p_3, width = 10, height = 10)

p_3
```


### Tech Users and Psych - Restless in diff Genders
```{r fig.width=10}
p_4 <- tech_mental_df |>
  group_by(YEAR, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Restless_1_percent = (Restless_1 / num_people),
    Restless_any_percent = (Restless_any / num_people)
  ) |>
  select(YEAR, Gender, Restless_1_percent,Restless_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Restless_1_percent", "Restless_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_wrap(~Gender_Str) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Restless") +
  ggtitle("Internet Use and Psychological Restlessness") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Restless_1_percent" = "lightblue","Restless_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Restless_1_percent" = "Restless All Time","Restless_any_percent" = "Restless Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_restless_gender.png", plot = p_4, width = 10, height = 10)

p_4

```


### Tech Users and Psych - Nervous in diff Genders and Age Group
```{r fig.width=10}
p_5 <- tech_mental_df |>
  group_by(YEAR, Age, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Nervous_1_percent = (Nervous_1 / num_people),
    Nervous_any_percent = (Nervous_any / num_people)
  ) |>
  select(YEAR, Age, Gender, Nervous_1_percent,Nervous_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Nervous_1_percent", "Nervous_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  mutate(
    Age_Group = case_when(
      Age==4 ~ "18-20 year olds",
      Age==5 ~ "21-25 year olds",
      Age==6 ~ "26-34 year olds",
      Age==7 ~ ">35 year olds"
      )
    ) |> 
  na.omit() |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_grid(Gender_Str ~ Age_Group) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Nervous ") +
  ggtitle("Internet Use and Psychological Nervousness by Age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Nervous_1_percent" = "lightblue","Nervous_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Nervous_1_percent" = "Nervous All Time","Nervous_any_percent" = "Nervous Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_nervous_gender_age.png", plot = p_5, width = 10, height = 10)

p_5
```


### Tech Users and Psych - Hopeless in diff Genders and Age Group
```{r fig.width=10}
p_6 <- tech_mental_df |>
  group_by(YEAR, Age, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
mutate(
    net_use_percent = (net_use/ n),
    Hopeless_1_percent = (Hopeless_1 / num_people),
    Hopeless_any_percent = (Hopeless_any / num_people)
  ) |>
  select(YEAR,Age, Gender, Hopeless_1_percent,Hopeless_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Hopeless_1_percent", "Hopeless_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  mutate(
    Age_Group = case_when(
      Age==4 ~ "18-20 year olds",
      Age==5 ~ "21-25 year olds",
      Age==6 ~ "26-34 year olds",
      Age==7 ~ ">35 year olds"
      )
    ) |> 
  na.omit() |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_grid(Gender_Str ~ Age_Group) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Hopeless") +
  ggtitle("Internet Use and Psychological Hopelessness by Age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Hopeless_1_percent" = "lightblue","Hopeless_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Hopeless_1_percent" = "Hopeless All Time","Hopeless_any_percent" = "Hopeless Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_hopeless_gender_age.png", plot = p_6, width = 10, height = 10)

p_6

```


### Tech Users and Psych - Restless in diff Genders and Age Group
```{r fig.width=10}
p_7 <- tech_mental_df |>
  group_by(YEAR, Age, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Restless_1_percent = (Restless_1 / num_people),
    Restless_any_percent = (Restless_any / num_people)
  ) |>
  select(YEAR, Age, Gender, Restless_1_percent,Restless_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Restless_1_percent", "Restless_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  mutate(
    Age_Group = case_when(
      Age==4 ~ "18-20 year olds",
      Age==5 ~ "21-25 year olds",
      Age==6 ~ "26-34 year olds",
      Age==7 ~ ">35 year olds"
      )
    ) |> 
  na.omit() |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group =
               as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_grid(Gender_Str ~ Age_Group) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Restless") +
  ggtitle("Internet Use and Psychological Restlessness by Age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Restless_1_percent" = "lightblue","Restless_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Restless_1_percent" = "Restless All Time","Restless_any_percent" = "Restless Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_restless_gender_age.png", plot = p_7, width = 10, height = 10)

p_7

```





### Tech Users and Psych - Nervous in diff Genders and Income Group
```{r fig.width=10}
p_8 <- tech_mental_df |>
  group_by(YEAR, Income, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Nervous_1_percent = (Nervous_1 / num_people),
    Nervous_any_percent = (Nervous_any / num_people)
  ) |>
  select(YEAR, Income, Gender, Nervous_1_percent,Nervous_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Nervous_1_percent", "Nervous_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  mutate(
    Income_Group = case_when(
      Income==1 ~ "Less than $20,000",
      Income==2 ~ "$20,000 - $49,999",
      Income==3 ~ "$50,000 - $74,999",
      Income==4 ~ "$75,000 or More"
      )
    ) |> 
  na.omit() |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_grid(Gender_Str ~ Income_Group) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Nervous ") +
  ggtitle("Internet Use and Psychological Nervousness by Income") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Nervous_1_percent" = "lightblue","Nervous_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Nervous_1_percent" = "Nervous All Time","Nervous_any_percent" = "Nervous Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_nervous_gender_income.png", plot = p_8, width = 10, height = 10)

p_8
```






### Tech Users and Psych - Hopeless in diff Genders and Income Group
```{r fig.width=10}
p_9 <- tech_mental_df |>
  group_by(YEAR, Income, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
mutate(
    net_use_percent = (net_use/ n),
    Hopeless_1_percent = (Hopeless_1 / num_people),
    Hopeless_any_percent = (Hopeless_any / num_people)
  ) |>
  select(YEAR,Income, Gender, Hopeless_1_percent,Hopeless_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Hopeless_1_percent", "Hopeless_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  mutate(
     Income_Group = case_when(
      Income==1 ~ "Less than $20,000",
      Income==2 ~ "$20,000 - $49,999",
      Income==3 ~ "$50,000 - $74,999",
      Income==4 ~ "$75,000 or More"
      )
    ) |> 
  na.omit() |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group = as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_grid(Gender_Str ~ Income_Group) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Hopeless") +
  ggtitle("Internet Use and Psychological Hopelessness by Income") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Hopeless_1_percent" = "lightblue","Hopeless_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Hopeless_1_percent" = "Hopeless All Time","Hopeless_any_percent" = "Hopeless Any Time", "net_use_percent" = "Average net use"))

ggsave("plots/p_tech_hopeless_gender_income.png", plot = p_9, width = 10, height = 10)

p_9
```




### Tech Users and Psych - Restless in diff Genders and Income Group
```{r fig.width=10}
p_10 <- tech_mental_df |>
  group_by(YEAR, Income, Gender) |>
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) |>
  ungroup() |>
  na.omit() |> 
  filter(YEAR<=2010) |> 
  mutate(
    net_use_percent = (net_use/ n),
    Restless_1_percent = (Restless_1 / num_people),
    Restless_any_percent = (Restless_any / num_people)
  ) |>
  select(YEAR, Income, Gender, Restless_1_percent,Restless_any_percent,  net_use_percent) |> 
  pivot_longer(
    cols = c("Restless_1_percent", "Restless_any_percent", "net_use_percent"),
    names_to = c("Type"),
    values_to = "Type_Value"
  ) |> 
  mutate(
    Gender_Str = case_when(
      Gender==1 ~ "MALE",
      Gender==2 ~ "FEMALE"
    )
  ) |> 
  mutate(
    Income_Group = case_when(
      Income==1 ~ "Less than $20,000",
      Income==2 ~ "$20,000 - $49,999",
      Income==3 ~ "$50,000 - $74,999",
      Income==4 ~ "$75,000 or More"
      )
    )  |> 
  na.omit() |> 
  ggplot(aes(x = as.factor(YEAR), y = Type_Value, color = as.factor(Type), group =
               as.factor(Type))) +
  geom_line(size = 1.2) +
  geom_point(size=2) +
  facet_grid(Gender_Str ~ Income_Group) +
  scale_y_continuous(labels = label_percent()) +
  xlab("Year") +
  ylab("% People Using Internet and Restless") +
  ggtitle("Internet Use and Psychological Restlessness by Income") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_colour_manual(name = "Attribute",
    values = c("Restless_1_percent" = "lightblue","Restless_any_percent" = "orange","net_use_percent" = "darkgreen"),
    labels = c("Restless_1_percent" = "Restless All Time","Restless_any_percent" = "Restless Any Time", "net_use_percent" = "Average net use"))


ggsave("plots/p_tech_restless_gender_income.png", plot = p_10, width = 10, height = 10)

p_10

```




