---
title: "Processing Attributes"
author: "Anurag Mallick"
date: "2025-04-28"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
fontsize: 12pt
classoption: twoside
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


```{r}
library(tidyverse)
library(stringr)
```






We use the codebook information to preprocess each attribute present and that can be used in the analysis.
At the end we create a final attribute file containing all the attributes of each year that we are considering.


`QUESTID2` is the identification variable. So we add it to each of our attribute pulls. `YEAR` we also add year to our attribute pulls.

The above two create the primary key for us.


## Data From All Attributes
```{r}
demo_vars_tib <- read_csv("variable_stats/Demographic_Vars_2002-19.csv", show_col_types = FALSE)
income_vars_tib <- read_csv("variable_stats/Income_Vars_2002-19.csv", show_col_types = FALSE)
edu_vars_tib <- read_csv("variable_stats/Education_Vars_2002-19.csv",show_col_types = FALSE)
mental_vars_tib <- read_csv("variable_stats/Mental_Health_Vars_2002-19.csv",show_col_types = FALSE)
psych_vars_tib <- read_csv("variable_stats/Psych_Distress_Vars_2002-19.csv",show_col_types = FALSE)
```

### Column Names
```{r}
demo_vars <- demo_vars_tib |> select(Column) |> unique()
demo_vars <- demo_vars[[1]]

income_vars <- income_vars_tib |> select(Column) |> unique()
income_vars <- income_vars[[1]]

edu_vars <- edu_vars_tib |> select(Column) |> unique()
edu_vars <- edu_vars[[1]]

mental_vars <- mental_vars_tib |> select(Column) |> unique()
mental_vars <- mental_vars[[1]]

psych_vars <- psych_vars_tib |> select(Column) |> unique()
psych_vars <- psych_vars[[1]]
```

```{r}
years <- demo_vars_tib |> select(Year) |> unique()
years <- years[[1]]
```

```{r}
all_req_vars <- c(demo_vars, income_vars, edu_vars, mental_vars, psych_vars)
```


### Read Data

We first get the data from 2002 till 2008. As many of the attributes is not uniquely available.
```{r}
final_data <- tibble()

years <- 2002:2007

for (year in years) {
  cat("\nFor year", year,"\n")
  
  
  file_name <- paste0("../health_data/raw/NSDUH_", year, "_Tab.txt")
  data <- read_tsv(file_name, col_types = cols(.default = "c"))
  data <- data |> select(all_of(c("QUESTID2",all_req_vars)))
  
  final_data <- bind_rows(final_data, data)
}
```

```{r}
# Storing the final data we use for visualization
# write_csv(final_data, "../health_data/data_2002-2007.csv")
```

```{r}
# Reading the final_data stored already
final_data <- read_csv("../health_data/processed/data_2002-2007.csv")
```


```{r}
final_data |> head(2)
```

## Process Data - with some EDA

## Demographics
```{r}
num_surveryed <- final_data |> group_by(YEAR) |> summarize(num_people=n()) |> ungroup()
```

### People Surveyed by Gender
1 = Male
2 = Female

```{r}
num_surveyed_gender <- final_data |>
  group_by(YEAR, IRSEX) |> 
  summarize(num_people=n()) |>
  ungroup()
  
num_surveyed_gender |> 
  pivot_wider(names_from = IRSEX, values_from = num_people, names_prefix = "Sex_")
```


### People Surveyed by Age

We use CATAG7 as the AGE attribute

1 = 12-13 Years Old
2 = 14-15 Years Old
3 = 16-17 Years Old
4 = 18-20 Years Old 
5 = 21-25 Years Old 
6 = 26-34 Years Old 
7 = 35 or Older


```{r}
num_surveyed_age <- final_data |> group_by(YEAR, CATAG7) |> 
  summarize(num_people=n()) |> 
  ungroup()

num_surveyed_age |> 
  pivot_wider(names_from = CATAG7, values_from = num_people, names_prefix = "AGE_")
```


### People Surveyed by Income

Given below is the median household income for different years
2002-01-01	42410
2003-01-01	43320
2004-01-01	44330
2005-01-01	46330
2006-01-01	48200
2007-01-01	50230

First - we use INCOME as the variable
1 = Less than $20,000 
2 = $20,000 - $49,999 
3 = $50,000 - $74,999 
4 = $75,000 or More 


```{r}
num_surveyed_income <- final_data |> group_by(YEAR, INCOME) |> 
  summarize(num_people=n()) |> 
  ungroup()

num_surveyed_income |> 
  pivot_wider(names_from = INCOME, values_from = num_people, names_prefix = "INCOME_")
```

We also use whether enrolled in government assist program or not as a variable

```{r}
num_surveyed_govtprog <- final_data |> group_by(YEAR, GOVTPROG) |> 
  summarize(num_people=n()) |> 
  ungroup()
```


### People Surveyed by Education Level


We use the variable EDUCCAT2
-9 = Data not collected for this year 
1 = Less than high school 
2 = High school graduate 
3 = Some college 
4 = College graduate  
5 = 12 to 17 year olds 

```{r}
num_surveyed_edu <- final_data |> group_by(YEAR, EDUCCAT2) |> 
  summarize(num_people=n()) |> 
  ungroup()

num_surveyed_edu |> 
  pivot_wider(names_from = EDUCCAT2, values_from = num_people, names_prefix = "EDU_")
```


### People Surveyed by All Levels
```{r}
num_surveyed_all_grp <- final_data |> group_by(YEAR, IRSEX, CATAG7, INCOME, EDUCCAT2) |>
  summarize(num_people=n()) |> 
  ungroup()
```


## Psych Distress Variables

### Number of Empty Variables
```{r}
final_data |> 
  select(YEAR, DSTNRVOS, DSTHOPLS, DSTRSTLS) |> 
  group_by(YEAR) |> 
  summarize(across(everything(),~sum(is.na(.x)), .names = "na_{.col}"))
```




### DSTNRVOS - HOW OFT FELT NERVOUS IN WORST MONTH, PST 12 MOS

-9 = Data not collected for this year 
1 = All of the time 
2 = Most of the time 
3 = Some of the time 
4 = A little of the time 
5 = None of the time 
85 = BAD DATA Logically assigned 
89 = LEGITIMATE SKIP Logically assigned 
94 = DON'T KNOW 
97 = REFUSED 
98 = BLANK (NO ANSWER) 
99 = LEGITIMATE SKIP 

```{r}
## Create a new variable for feeling nervous 
## 1,2,3,4 - Any time felt nervous - "_any"
## 5 - No time felt nervous - "_no"
## 99 - Legitimate Skip - "_skip"
## Others - "_others"
## -9 - "_notcollect"

final_data <- final_data |>
  mutate(DSTNRVOS_enc = case_when(
    DSTNRVOS %in% 1:4  ~ "any",
    DSTNRVOS == 5      ~ "no",
    DSTNRVOS == 99     ~ "skip",
    DSTNRVOS == -9     ~ "notcollect",
    TRUE                  ~ "others"
  ))

```


```{r, message = FALSE, warning = FALSE}
# Creating overall summary of nervous people
nervous_summ <- final_data |> group_by(YEAR, DSTNRVOS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_enc <- final_data |> group_by(YEAR, DSTNRVOS_enc) |> 
  summarize(num_people=n()) |>
  pivot_wider(names_from = DSTNRVOS_enc, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_ovrl <- nervous_summ |> left_join(nervous_summ_enc)
nervous_summ_ovrl <- nervous_summ_ovrl |> left_join(num_surveryed) |> ungroup()
```


```{r}
# Plotting the nervousness in people over the years
nervous_summ_ovrl |> 
  mutate(Nervous_any_percent=Nervous_any/num_people) |> 
  ggplot(mapping = aes(x=YEAR, y=Nervous_any_percent)) +
  geom_line(group=1, size=1.5)
```


#### By Gender Group
```{r, message=FALSE}
nervous_summ_sex <- final_data |> 
  group_by(YEAR,IRSEX, DSTNRVOS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_sex_enc <- final_data |> 
  group_by(YEAR,IRSEX, DSTNRVOS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS_enc, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_sex_ovrl <- nervous_summ_sex |> left_join(nervous_summ_sex_enc, by=c("YEAR","IRSEX"))
nervous_summ_sex_ovrl <- nervous_summ_sex_ovrl |> left_join(num_surveyed_gender, by=c("YEAR","IRSEX"))
```


#### By Age Group
```{r, message=FALSE}
nervous_summ_age <- final_data |> 
  group_by(YEAR,CATAG7, DSTNRVOS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS, values_from = num_people, names_prefix = "Nervous_") |> 
  filter(CATAG7%in%c(4,5,6,7)) |> 
  ungroup()

nervous_summ_age_enc <- final_data |> 
  group_by(YEAR,CATAG7, DSTNRVOS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS_enc, values_from = num_people, names_prefix = "Nervous_") |> 
  filter(CATAG7%in%c(4,5,6,7)) |> 
  ungroup()

nervous_summ_age_ovrl <- nervous_summ_age |> left_join(nervous_summ_age_enc, by=c("YEAR","CATAG7"))
nervous_summ_age_ovrl <- nervous_summ_age_ovrl |> left_join(num_surveyed_age, by=c("YEAR","CATAG7"))
```


```{r}
## Plotting the nervousness by different age groups
nervous_summ_age_ovrl |> 
  mutate(Nervous_any_percent=Nervous_any/num_people) |> 
  ggplot(mapping = aes(x=YEAR, y=Nervous_any_percent)) +
  geom_line(aes(group = CATAG7, colour = CATAG7), size=1.5)
```

```{r}
## Plotting the nervousness by different age groups
nervous_summ_age_ovrl |> 
  mutate(Nervous_1_percent=Nervous_1/num_people) |> 
  ggplot(mapping = aes(x=YEAR, y=Nervous_1_percent)) +
  geom_line(aes(group = CATAG7, colour = CATAG7), size=1.5)
```



#### By Income Group
```{r}
nervous_summ_income <- final_data |> 
  group_by(YEAR,INCOME, DSTNRVOS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_income_enc <- final_data |> 
  group_by(YEAR,INCOME, DSTNRVOS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS_enc, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_income_ovrl <- nervous_summ_income |> left_join(nervous_summ_income_enc, by=c("YEAR","INCOME"))
nervous_summ_income_ovrl <- nervous_summ_income_ovrl |> left_join(num_surveyed_income, by=c("YEAR","INCOME"))
```


```{r}
## Plotting the nervousness by different age groups
nervous_summ_income_ovrl |> 
  mutate(Nervous_any_percent=Nervous_any/num_people) |> 
  ggplot(mapping = aes(x=YEAR, y=Nervous_any_percent)) +
  geom_line(aes(group = INCOME, colour = INCOME), size=1.5)
```

```{r}
## Plotting the nervousness by different age groups
nervous_summ_income_ovrl |> 
  mutate(Nervous_1_percent=Nervous_1/num_people) |> 
  ggplot(mapping = aes(x=YEAR, y=Nervous_1_percent)) +
  geom_line(aes(group = INCOME, colour = INCOME), size=1.5)
```



#### By Education Group
```{r}
nervous_summ_edu <- final_data |> 
  group_by(YEAR,EDUCCAT2, DSTNRVOS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_edu_enc <- final_data |> 
  group_by(YEAR,EDUCCAT2, DSTNRVOS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS_enc, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_edu_ovrl <- nervous_summ_edu |> left_join(nervous_summ_edu_enc, by=c("YEAR","EDUCCAT2"))
nervous_summ_edu_ovrl <- nervous_summ_edu_ovrl |> left_join(num_surveyed_edu, by=c("YEAR","EDUCCAT2"))
```


#### By All Demographic Groups
```{r}
nervous_summ_all_grp <- final_data |> 
  group_by(YEAR, IRSEX, CATAG7, INCOME, EDUCCAT2, DSTNRVOS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_all_grp_enc <- final_data |> 
  group_by(YEAR, IRSEX, CATAG7, INCOME, EDUCCAT2,  DSTNRVOS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTNRVOS_enc, values_from = num_people, names_prefix = "Nervous_") |> 
  ungroup()

nervous_summ_all_grp_ovrl <- nervous_summ_all_grp |> left_join(nervous_summ_all_grp_enc, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2"))
nervous_summ_all_grp_ovrl <- nervous_summ_all_grp_ovrl |> left_join(num_surveyed_all_grp, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2"))
```








### DSTHOPLS - HOW OFTEN FELT HOPELESS IN WORST MONTH

-9 = Data not collected for this year 
1 = All of the time 
2 = Most of the time 
3 = Some of the time 
4 = A little of the time 
5 = None of the time 
85 = BAD DATA Logically assigned 
89 = LEGITIMATE SKIP Logically assigned 
94 = DON'T KNOW 
97 = REFUSED 
98 = BLANK (NO ANSWER) 
99 = LEGITIMATE SKIP 


```{r}
## Create a new variable for feeling hopeless 
## 1,2,3,4 - Any time felt hopeless - "_any"
## 5 - No time felt hopeless - "_no"
## 99 - Legitimate Skip - "_skip"
## Others - "_others"
## -9 - "_notcollect"

final_data <- final_data |>
  mutate(DSTHOPLS_enc = case_when(
    DSTHOPLS %in% 1:4  ~ "any",
    DSTHOPLS == 5      ~ "no",
    DSTHOPLS == 99     ~ "skip",
    DSTHOPLS == -9     ~ "notcollect",
    TRUE                  ~ "others"
  ))

```


```{r}
hopeless_summ <- final_data |> group_by(YEAR, DSTHOPLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_enc <- final_data |> group_by(YEAR, DSTHOPLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS_enc, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()


hopeless_summ_ovrl <- hopeless_summ |> left_join(hopeless_summ_enc)
hopeless_summ_ovrl <- hopeless_summ_ovrl |> left_join(num_surveryed) |> ungroup()
```



#### By Gender Group
```{r, message=FALSE}
hopeless_summ_sex <- final_data |> 
  group_by(YEAR,IRSEX, DSTHOPLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_sex_enc <- final_data |> 
  group_by(YEAR,IRSEX, DSTHOPLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS_enc, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_sex_ovrl <- hopeless_summ_sex |> left_join(hopeless_summ_sex_enc, by=c("YEAR","IRSEX"))
hopeless_summ_sex_ovrl <- hopeless_summ_sex_ovrl |> left_join(num_surveyed_gender, by=c("YEAR","IRSEX"))
```



#### By Age Group
```{r}
hopeless_summ_age <- final_data |> 
  group_by(YEAR,CATAG7, DSTHOPLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS, values_from = num_people, names_prefix = "Hopeless_") |> 
  filter(CATAG7%in%c(4,5,6,7)) |> 
  ungroup()

hopeless_summ_age_enc <- final_data |> 
  group_by(YEAR,CATAG7, DSTHOPLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS_enc, values_from = num_people, names_prefix = "Hopeless_") |> 
  filter(CATAG7%in%c(4,5,6,7)) |> 
  ungroup()

hopeless_summ_age_ovrl <- hopeless_summ_age |> left_join(hopeless_summ_age_enc, by=c("YEAR","CATAG7"))
hopeless_summ_age_ovrl <- hopeless_summ_age_ovrl |> left_join(num_surveyed_age, by=c("YEAR","CATAG7"))
```



#### By Income Group
```{r}
hopeless_summ_income <- final_data |> 
  group_by(YEAR,INCOME, DSTHOPLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_income_enc <- final_data |> 
  group_by(YEAR,INCOME, DSTHOPLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS_enc, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_income_ovrl <- hopeless_summ_income |> left_join(hopeless_summ_income_enc, by=c("YEAR","INCOME"))
hopeless_summ_income_ovrl <- hopeless_summ_income_ovrl |> left_join(num_surveyed_income, by=c("YEAR","INCOME"))
```



#### By Education Group
```{r}
hopeless_summ_edu <- final_data |> 
  group_by(YEAR,EDUCCAT2, DSTHOPLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_edu_enc <- final_data |> 
  group_by(YEAR,EDUCCAT2, DSTHOPLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS_enc, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_edu_ovrl <- hopeless_summ_edu |> left_join(hopeless_summ_edu_enc, by=c("YEAR","EDUCCAT2"))
hopeless_summ_edu_ovrl <- hopeless_summ_edu_ovrl |> left_join(num_surveyed_edu, by=c("YEAR","EDUCCAT2"))
```


#### By All Demographic Groups
```{r}
hopeless_summ_all_grp <- final_data |> 
  group_by(YEAR, IRSEX, CATAG7, INCOME, EDUCCAT2, DSTHOPLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_all_grp_enc <- final_data |> 
  group_by(YEAR, IRSEX, CATAG7, INCOME, EDUCCAT2,  DSTHOPLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTHOPLS_enc, values_from = num_people, names_prefix = "Hopeless_") |> 
  ungroup()

hopeless_summ_all_grp_ovrl <- hopeless_summ_all_grp |> left_join(hopeless_summ_all_grp_enc, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2"))
hopeless_summ_all_grp_ovrl <- hopeless_summ_all_grp_ovrl |> left_join(num_surveyed_all_grp, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2"))
```














### DSTRSTLS - HOW OFTEN FELT restless IN WORST MONTH

-9 = Data not collected for this year 
1 = All of the time 
2 = Most of the time 
3 = Some of the time 
4 = A little of the time 
5 = None of the time 
85 = BAD DATA Logically assigned 
89 = LEGITIMATE SKIP Logically assigned 
94 = DON'T KNOW 
97 = REFUSED 
98 = BLANK (NO ANSWER) 
99 = LEGITIMATE SKIP 


```{r}
## Create a new variable for feeling restless 
## 1,2,3,4 - Any time felt restless - "_any"
## 5 - No time felt restless - "_no"
## 99 - Legitimate Skip - "_skip"
## Others - "_others"
## -9 - "_notcollect"

final_data <- final_data |>
  mutate(DSTRSTLS_enc = case_when(
    DSTRSTLS %in% 1:4  ~ "any",
    DSTRSTLS == 5      ~ "no",
    DSTRSTLS == 99     ~ "skip",
    DSTRSTLS == -9     ~ "notcollect",
    TRUE                  ~ "others"
  ))

```


```{r}
restless_summ <- final_data |> group_by(YEAR, DSTRSTLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_enc <- final_data |> group_by(YEAR, DSTRSTLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS_enc, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()


restless_summ_ovrl <- restless_summ |> left_join(restless_summ_enc)
restless_summ_ovrl <- restless_summ_ovrl |> left_join(num_surveryed) |> ungroup()
```



#### By Gender Group
```{r, message=FALSE}
restless_summ_sex <- final_data |> 
  group_by(YEAR,IRSEX, DSTRSTLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_sex_enc <- final_data |> 
  group_by(YEAR,IRSEX, DSTRSTLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS_enc, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_sex_ovrl <- restless_summ_sex |> left_join(restless_summ_sex_enc, by=c("YEAR","IRSEX"))
restless_summ_sex_ovrl <- restless_summ_sex_ovrl |> left_join(num_surveyed_gender, by=c("YEAR","IRSEX"))
```


#### By Age Group
```{r}
restless_summ_age <- final_data |> 
  group_by(YEAR,CATAG7, DSTRSTLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS, values_from = num_people, names_prefix = "Restless_") |> 
  filter(CATAG7%in%c(4,5,6,7)) |> 
  ungroup()

restless_summ_age_enc <- final_data |> 
  group_by(YEAR,CATAG7, DSTRSTLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS_enc, values_from = num_people, names_prefix = "Restless_") |> 
  filter(CATAG7%in%c(4,5,6,7)) |> 
  ungroup()

restless_summ_age_ovrl <- restless_summ_age |> left_join(restless_summ_age_enc, by=c("YEAR","CATAG7"))
restless_summ_age_ovrl <- restless_summ_age_ovrl |> left_join(num_surveyed_age, by=c("YEAR","CATAG7"))
```



#### By Income Group
```{r}
restless_summ_income <- final_data |> 
  group_by(YEAR,INCOME, DSTRSTLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_income_enc <- final_data |> 
  group_by(YEAR,INCOME, DSTRSTLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS_enc, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_income_ovrl <- restless_summ_income |> left_join(restless_summ_income_enc, by=c("YEAR","INCOME"))
restless_summ_income_ovrl <- restless_summ_income_ovrl |> left_join(num_surveyed_income, by=c("YEAR","INCOME"))
```



#### By Education Group
```{r}
restless_summ_edu <- final_data |> 
  group_by(YEAR,EDUCCAT2, DSTRSTLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_edu_enc <- final_data |> 
  group_by(YEAR,EDUCCAT2, DSTRSTLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS_enc, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_edu_ovrl <- restless_summ_edu |> left_join(restless_summ_edu_enc, by=c("YEAR","EDUCCAT2"))
restless_summ_edu_ovrl <- restless_summ_edu_ovrl |> left_join(num_surveyed_edu, by=c("YEAR","EDUCCAT2"))
```




#### By All Demographic Groups
```{r}
restless_summ_all_grp <- final_data |> 
  group_by(YEAR, IRSEX, CATAG7, INCOME, EDUCCAT2, DSTRSTLS) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_all_grp_enc <- final_data |> 
  group_by(YEAR, IRSEX, CATAG7, INCOME, EDUCCAT2,  DSTRSTLS_enc) |> 
  summarize(num_people=n()) |> 
  pivot_wider(names_from = DSTRSTLS_enc, values_from = num_people, names_prefix = "Restless_") |> 
  ungroup()

restless_summ_all_grp_ovrl <- restless_summ_all_grp |> left_join(restless_summ_all_grp_enc, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2"))
restless_summ_all_grp_ovrl <- restless_summ_all_grp_ovrl |> left_join(num_surveyed_all_grp, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2"))
```




### Merging All Overall Attributes

```{r}
# nervous_summ_ovrl - Feeling of nervousness overall
# hopeless_summ_ovrl - Feeling of hopelessness overall 
# restless_summ_ovrl - Feeling of restlessness overall
```

```{r}
psych_summ_ovrl <- nervous_summ_ovrl |> 
  left_join(restless_summ_ovrl) |> 
  left_join(hopeless_summ_ovrl)

# write_csv(psych_summ_ovrl,"variable_eda/psych_summ_ovrl.csv")
```


### Merging All Gender Group Attributes
```{r}
psych_summ_sex_ovrl <- nervous_summ_sex_ovrl |> 
  left_join(hopeless_summ_sex_ovrl, by=c("YEAR", "IRSEX", "num_people")) |>
  left_join(restless_summ_sex_ovrl, by=c("YEAR", "IRSEX", "num_people"))

# write_csv(psych_summ_sex_ovrl,"variable_eda/psych_summ_sex_ovrl.csv")
```



### Merging All Age Group Attributes
```{r}
psych_summ_age_ovrl <- nervous_summ_age_ovrl |> 
  left_join(hopeless_summ_age_ovrl, by=c("YEAR", "CATAG7", "num_people")) |>
  left_join(restless_summ_age_ovrl, by=c("YEAR", "CATAG7", "num_people"))

# write_csv(psych_summ_age_ovrl,"variable_eda/psych_summ_age_ovrl.csv")
```


### Merging All Income Group Attributes
```{r}
psych_summ_income_ovrl <- nervous_summ_income_ovrl |>
  left_join(hopeless_summ_income_ovrl, by=c("YEAR", "INCOME", "num_people")) |>
  left_join(restless_summ_income_ovrl, by=c("YEAR", "INCOME", "num_people"))

# write_csv(psych_summ_income_ovrl,"variable_eda/psych_summ_income_ovrl.csv")
```


### Merging All Education Group Attributes
```{r}
psych_summ_edu_ovrl <- nervous_summ_edu_ovrl |> 
  left_join(hopeless_summ_edu_ovrl, by=c("YEAR", "EDUCCAT2", "num_people")) |>left_join(restless_summ_edu_ovrl, by=c("YEAR", "EDUCCAT2", "num_people"))

write_csv(psych_summ_edu_ovrl,"variable_eda/psych_summ_edu_ovrl.csv")
```



### Merging All DEMOGRAPHIC Group Attributes
```{r}
psych_summ_all_grp_ovrl <- nervous_summ_all_grp_ovrl |> 
  left_join(hopeless_summ_all_grp_ovrl, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2", "num_people")) |> 
  left_join(restless_summ_all_grp_ovrl, by=c("YEAR","IRSEX", "CATAG7", "INCOME", "EDUCCAT2", "num_people"))

write_csv(psych_summ_all_grp_ovrl,"variable_eda/psych_summ_all_grp_ovrl.csv")
```

